### 2019-01-13 Write_up/Writing/Introduction_outline.Rmd
Introduction outline

---
title: "Introduction_outline"
author: "Amieroh Abrahams"
date: "13 January 2019"
output: html_document
---

### Introduction

Seawater temperature is a key indicator of environmental change in both terrestrial and marine ecosystems (Pearce et al., 2006), which directly influences the earth’s natural ecological functioning. Climate change is widely believed to result in a rise in the earth’s average surface temperatures over time (Stocker et al., 2013) and this is usually associated with an increase in the intensity and duration of extreme weather events such as cold-spells, heatwaves, floods and storms. Consequently, these extreme climatic events can often have profound short-term effects on both humans as well as many ecosystems (Easterling et al., 2000).

‘Cold spells’ are usually referred to as a phenomenon where abnormally cold water remains on the surface of the ocean for extended periods of time. However, several other definitions exist based on the statistical properties of the temperature record in association with the location of temperature collections and the time of year .  Studies that have quantified the impacts of extreme temperatures have provided evidence of demonstrably negative effects on  associated surrounding regions as well as the specific regions within which these climatic events occur. While extensive research has been done on other extreme weather events such as floods, wind storms and marine heat waves, minimal research has been done on marine cold spells (MCS). It is known that warm-water anomalies may result in severe coral bleaching and is one of the factors resulting in the demise of coral reefs. Additionally, anomalous warm water events are known to have negative impacts on species diversity and mortality as well as the distribution of organisms. However, recent studies done by Lirman et al. explains that corals living in regions or surrounding regions of unusually prolonged anonymously cold waters may be just as detrimental; this was illustrated in January of 2010 when the reefs of the Florida Reef Track (FRT) were exposed to an unusual cold water event that resulted in large scale coral mortality. This climatic event showed clear association with North Atlantic Oscillation (NAO) index that produced northerly surface wind anomalies and the southward advection of the cold Arctic air. Similarly, research by Shinn et al. (1962) reported high mortality of Acropora cervicornis in shallow reefs as temperatures dropped. Likewise, studies conducted by Hudson et al.  also documented the mortality of Montastraea colonies in near shore patch reefs due to cold fronts between 1969–1970 and 1977–1978, respectively.

MCSs are lethal to all marine organisms and are known to have caused  mass extinctions of several fish and invertebrate species (Woodward. 1987 & Gunter. 1941, 1951 & Holt and Holt. 1983, Gunter. 1951 & Crisp. 1964). Temperature has long been considered one of the primary factors controlling species distribution patterns by limiting their distribution north or southward towards the high latitudes (Firth et al., 2011). Studies done by Firth et al. (2011, 2015) have demonstrated the range shifts of engineer species such as mussels as a result of MCSs. Thus, providing a clearer understanding of the influence of oceanic and atmospheric processes on sea surface temperatures (SST) within the upper mixed layer, however minimal research has been done on the influences of local and broad scale impacts on seawater temperature.

Research by Grunter (1941) and Firth et al. (2011) explains that some MCSs that are believed to have influenced species populations were caused by atmospheric cold spells which had ultimately affected the local intertidal and coastal biota. MCSs have been proven to have a great effect on ecosystem change, thus understanding these climatic drivers may provide useful tools for conservation and management efforts. This study will aim to develop a better understanding of the impacts of weak and strong upwelling events and how they may be influencing the intensity and duration of MCSs. This study applied the MCSs definitions to datasets at a local-scale and an offshore broad scale.

Eastern Boundary Upwelling Systems (EBUS) are among the most productive regions of the world. These upwelling systems include the; California, Humboldt, Canary and Benguela currents, and are important for biological productivities by supporting commercial fisheries (Pauly and Christensen, 1995) thus forming an essential component to these region’s economies, as well as providing recreational services to the human population along their coasts. One of the five continental margin upwelling systems is the Benguela Current; an eastern boundary current of the south Atlantic subtropical gyre which flows towards the equator approximately 300 km offshore of the coast of southwest Africa before splitting into the Benguela Coastal Current (BCC) and Benguela Oceanic Current (BOC). Shoreward of the main stream of the Benguela current there is a slow flow along the continental margin into the Angola Basin. In this slow flow the prevailing south and south eastern winds drive coastal upwelling of cold, nutrient rich South Atlantic Central Water (SACW).

The SACW originates by sinking at the Subtropical Front, known as the Subtropical Convergence (Schell, 1970 & Nelson, 1989). Beneath the SACW lies cold, nutrient-rich Antarctic Intermediate Water (AAIW) which originates by sinking at the Polar Front. Throughout each of the seasons, upwelling of nutrient rich waters covers most of the continental shelf and much of the continental slope (Shannon, 1985), thus creating what we refer to as the Benguela Current Upwelling System. Research studies have shown that these cold waters move north over the shelf in the Benguela Coastal Current (Hay and Brock, 1992). Offshore they are separated from the warmer waters of the subtropical Atlantic by a well-developed thermal front (Shannon, 1985). This quasistationary region is often linked with the shelf edge, although it is distorted by eddies, plumes and jets which extend west into the SE Atlantic (Lutmharms and Stockton, 1987; Waldron and Probyn, 1992). These plumes largely expand the region of high productivity as a result of upwelling by rapidly advecting cold, nutrient-rich water from the coastal upwelling sites into the open ocean (Lutjeiarms et al, 1991). Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current (Peterson and Stramma, 1991). Southward, the coastal upwelling system is bounded by the Agulhas Bank. South of the African coast the warm Agulhas Current flows in from the Indian Ocean and returns eastwards while generating warm eddies that spin off in a north-westerly direction in the Benguela current (Duncombe, 1991). Research studies done by Shannon et al. (1985) explains that coastal upwelling varies on a seasonal basis and that during winter and spring, water of approximately 16°C extends along the coast. Upwelling is coldest and most persistent off the Namibian coast in the Lüderitz region at 25°S and weakest at 17°S (Lutjeharms and Meeuwis, 1987). Lutjeharms and Meeuwis, (1987) also found a strong relationship between the intensity of upwelling events and the strength and direction of prevailing winds.  It is also hypothesised that upwelling is more intense where deep water is closer to the coast.

EBUS mesoscale and sub-mesoscale oceanic movements are known to influence marine ecosystems in that it enhances primary production (McGillicuddy et al., 1998; Oschlies and Garçon, 1998) and is key in managing the exchange and of transporting of organic resources and marine organisms from the coastal zone to the open ocean. Large-scale atmospheric pressure systems that favour along-shore, equatorward winds, result in the high levels of productivity in the EBUS, which in combination with the Coriolis effect, affects surface water offshore. This results in a process whereby cold, nutrient rich, with low pH and low oxygen concentration, originates from deep within the ocean and is pushed to the surface where they, in the form of phytoplankton blooms, fuel primary production. As a result of the intensity and duration of the upwelling event; regions affected by these cold water events are exposed to distinct ecological conditions and tend to differ from its surrounding regions (Bosman et al., 1987; Barshis et al., 2011). Significant research evidence suggests that SST have increased, however some regions still show a lower or non-significant increase, and in some cases cooling trends are evident. Inshore regions and embayment’s that are protected from wind action may be subjected to intense episodes of primary production (Graham and Largier, 1997). Vander Wouder et al. (2006) found that these sites represent the highest levels of chlorophyll concentration.

Atmospheric and oceanographic processes results in upwelling, while the amplitude and timing of upwelling-favourable winds are sensitive to climate variability (Montecinos et al., 2003; Macias et al., 2012), and are highly likely to be affected by global warming (Barros et al., 2014; Bakun et al., 2015). It is hypothesised that global warming will influence atmospheric pressure gradients, ultimately affecting the amplitude and timing of coastal winds that cause upwelling (Bakun, 1990; Bakun et al., 2015) and more recent observations (Rykaczewski et al., 2015; Wanget al., 2015) by means of global model projections strongly suggests intensification of coastal winds. Understanding the influence of changing winds on upwelling events is highly important. For example, weak winds may result in weak upwelling, which in turn may limit nutrient concentration; resulting in potentially negative impacts on primary production (Chhak and Di Lorenzo, 2007). In contrast, stronger upwelling winds, could theoretically result in stronger upwelling events, thereby increasing the nutrient concentration which may increase offshore transport (Cury and Roy, 1989). Increased wind action may also result in increased turbulence (Cury and Roy, 1989); and may even affect the processes associated with ocean acidification, creating widespread changes within ecosystems and ultimately impacting regional productivity (Gruber, 2011; Doney et al. 2012).

**** A better understanding of the impact of global warming on wind variability is important in that it allows for better management of fisheries and other marine resources in the EBUS which could be improved with a better understanding of how nutrient abundance and primary production are associated with upwelling events. This study aims to emphasize the drivers of upwelling-favourable winds, as well as the factors influencing nutrient enrichment, and potential changes in biogeochemistry*****

The Benguela current, forming part of the EBUS, experiences cross-shore atmospheric pressure gradients, which results in alongshore, equatorial winds. These winds are known to drive coastal upwelling events. With this in mind, research by Bakun (1990) suggests that an increase in greenhouse gas emissions would exacerbate warm-season upwelling. The understanding behind this mechanism was based on the idea that air above land would warm rapidly as compared to air above the ocean, and that the deepening of the continental thermal low pressure system (CTLPS) would increase the cross-shore pressure gradient which drives upwelling-favourable winds. Several research studies examining the trend between upwelling events and wind patterns, have however indicated contrasting results. With more than 20 years of observational and modelled data; Sydeman et al. (2014a) discovered the most coherent pattern of intensifying upwelling-favourable winds in the Humboldt and California upwelling systems. It has proven difficult to attribute coastal wind intensification in EBUS to global warming, despite the fact that the meta-analysis done by Sydeman et al. (2014a) generally supports Bakun’s (1990) proposition.

This study cannot discount the role of multi-decadal climate variability in the observed trends (Chhak and Di Lorenzo, 2007; Narayan et al., 2010; Pérez et al., 2010; Macias et al., 2012; Santos et al., 2012; Cropper et al., 2014; Jacox et al., 2014), nor is it a test of the Bakun (1990) hypothesis mechanism. Given the broad range of contradicting studies, the only conclusion that can be made is that upwelling-favourable winds tend to be the most intense during the warmer months of the year (Huyer, 1983; Nelson and Hutchings, 1983). 
The timing and the intensity of upwelling favourable winds are highly variable within the EBUS, therefore creating much uncertainty (Chavez and Messié, 2009). Understanding the factors contributing to the intensity and duration of upwelling events is important because variations in the timing of upwelling strongly influences ecosystem productivity. Since the late 1940s studies have shown an increase in upwelling favourable winds associated with the California current, however, less variability has been observed in the strength of upwelling winds through the year in the central Benguela and Canary Currents (Shannon and Nelson, 1996; Varela et al., 2015). It is important to note that the lack of research conducted may obscure the validity of any conclusion drawn here. This study further researched the changes in the timing and variability of upwelling within the Benguela Current and how these phenomena influence associated ecosystems.

As a result of climate change, global ocean temperatures have increased at an accelerating rate (Stocker et al., 2013). Sufficient research suggests that coastal sea surface temperatures (SST) in EBUS such as the Benguela current are cooler than offshore regions as a result of the near shore upwelling waters. There is a large but contradicting amount of evidence on the SST trends in EBUS (Mendelssohn and Schwing, 2002; Arístegui et al., 2009; Belkin, 2009; Demarcq, 2009; Lebassi et al., 2009; García-Reyes and Largier, 2010; Rouault et al., 2010; Gutiérrez et al., 2011; Seo et al., 2012; Barton et al., 2013; Barros et al., 2014; Sydeman et al., 2014b; Salvanes et al., 2015). The contradictions amongst studies exist because of varying resolutions of SST data used, thus resulting in the capturing of different aspects of the upwelling processes, showing different trends that exist based on the resolutions of the different SST datasets as well as other intrinsic factors. Generally, because of the resolution of SST datasets it is often difficult to separate near shore upwelling temperatures from regional, broad scale and eastern boundary current temperatures. Lima and Wethey’s (2012) found that high resolution data represented clear cooling trends predominantly during summer and at high latitudes in most of the EBUSs. Additionally several research studies examined the difference between coastal and offshore temperature trends, revealing decreasing or increasing trends in near shore SST, compared with increasing trends offshore in the Benguela compared with increasing trends offshore in the Benguela (Rouault et al.,2010; Santos et al., 2012; Salvanes et al., 2015). In this study however, will utilise a range of spatio-temporal resolutions of SST data and in situ collected seawater temperature measurements for the Benguela upwelling system. The aim of this is to demonstrate the variable upwelling events based on the different resolutions at which temperature is being collected.

Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends (Demarcq,2009). It is predicted that these climate oscillations may increase the variance with climate change (Timmermann et al., 1999; Kuzmina et al., 2005; Sydeman et al., 2013; Cai et al., 2015) which will further result in unusual trends.

Research studies provide significant evidence on the increase in warming of the ocean surface. This warming slowly propagates down through the water column and in turn increases the thermal stratification (Gruber, 2011). Strong stratification limits the depth at which water is up welled and hence controls the amount of nutrients brought up (Chhak and Di Lorenzo, 2007; Jacox and Edwards, 2011; Jacox et al., 2015). The interplay between upwelling winds and stratification determine the amount of nutrients present within the system and thus is strongly linked to biological productivity. Globally; as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% during the period 1971 to 2010 (Stocker et al., 2013). Variability in the biogeochemical properties of up welled waters influences species distribution, ecosystem structure and productivity within the Benguela upwelling system. The deep waters being pushed to the surface through the upwelling process is high in nutrients and dissolved CO2, a consequence of the long period of isolation at a depth from the euphotic zone (Feely et al., 2008). Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses. As the deep water is being up welled to the surface, decreases in oxygen and pH levels are predicted along with an increase in nutrient levels (Rykaczewski and Dunne, 2010).

Biotic systems within the Benguela upwelling system may be negatively influenced by the changes in oxygen levels given their adaption to natural variable conditions (Capone and Hutchins, 2013).  This study will seek to examine the impact of both strong and weak upwelling winds on the intensity and duration of the upwelling event; additionally it will also aim to examine how chlorophyll concentrations vary with regard to contrasting upwelling conditions as well as the impact this has on species distribution and habitats.

##### Ideas #####

-chnaging the structure it to describe MCS, explain the dangers of these phenomena on marine life (, the implications of this (eg distributions being affected), and the causes of MCS.

-explain what info is unknown about MCS and why this is a problem in areas where they can occur

- link this to the EBUS and Benguela stuff

- clearly mention the aims, objectives and approach to the study.