### 2019-02-12 wavelets.Rmd
updated

---
title: "Wavelets"
author: "Amieroh Abrahams"
date: "09 February 2019"
output: html_document
---

Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events varies. It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions. Here i was only required to plot a time series of 1year.

# Libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
```

### Load site list SACTN data
Now to get to the data. The first step involves the loading of the site list. The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures. These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents. Here we will only focus on the temperature data found along the west coast (wc) (i.e. sites influenced by the Benguela current, EBUS).

```{r}
load("Data/site_list_v4.2.RData")
load("Data/SACTN_daily_v4.2.RData")
```

## Load in the rest of the datasets

* See the Dectect_MCS.Rmd script for more info. The datasets being loaded here are only for temperatures collected between the years 1992-2016. However if a longer time series is required changes may  be made to the Dectect_MCS.Rmd script (line 136)

```{r}
load("~/Documents/Masters_2019/MastersProject/Data/overlap_ts.RData")
load("~/Documents/Masters_2019/MastersProject/Data/insitu_MUR.RData")
load("~/Documents/Masters_2019/MastersProject/Data/insitu_OISST.RData")
load("~/Documents/Masters_2019/MastersProject/Data/insitu_G1SST.RData")
load("~/Documents/Masters_2019/MastersProject/Data/insitu_K10.RData")
```

# Time series of one year

Here I only look at upwelling duration over a period of 1 year. from 2014 - 2015

```{r}
one_year <- function(df){
  upwelling<- df %>% 
  filter(year(date) %in% seq(2014, 2015)) %>% 
  drop_na()
}

SACTN_1yr_US1 <- one_year(df = overlap_ts) 
MUR_1yr_US1 <- one_year(df = insitu_MUR) 
OISST_1yr_US1 <- one_year(df = insitu_OISST) 
G1SST_1yr_US1 <- one_year(df = insitu_G1SST) 
K10_1yr_US1 <- one_year(df = insitu_K10) 
```

# Wavelet analyses

```{r}
temp.d <- function(df){
temp.d <- df %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, temp, date)
}

o
temp.d <- temp.d(df = overlap_ts) 
# temp.d <- temp.d(df = insitu_MUR) 
# temp.d <- temp.d(df = insitu_OISST) 
# temp.d <- temp.d(df = insitu_G1SST) 
# temp.d <- temp.d(df = insitu_K10)

temp.d <- temp.d %>% 
  select(site, no, temp, date)

prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))

ggplot(PN_prewhite, aes(x = no, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()

wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}

PN_wave <- dlply(PN_prewhite, .(site), wl.fun)

for (i in 1:length(PN_wave)) {
  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures",plot_name,"_wavelet.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(PN_wave, plot_fun)
```