### 2019-02-19 sites_metadata.Rmd
study_site

---
title: "sites_metadata"
author: "Amieroh Abrahams"
date: "19 February 2019"
output: html_document
---

```{r}
library(tidyverse)
library(viridis)
library(plyr)
library(zoo)
library(lubridate)
library(maptools)
library(sp)
library(geosphere)
library(marmap)
library(PBSmapping)
library(scales)
library(grid)
library(gridExtra)
library(mapproj)
```

```{r}
load("~/Documents/Masters_2019/MastersProject/Data/site_list_v4.2.RData")

# # Upwelling Cell 1:
# site_list_sub <- site_list %>%
#   filter(coast == "wc") %>%
#   filter(length > 3650)
#
# # upwelling_sites <- site_list_sub[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS
# upwelling_sites <- upwelling_sites[-10:-14,]
#
# # # # Here I find a time series with the data with a relativley long time series
# UC_1 <- upwelling_sites[c(-2, -3, -5, -6, -8),]
# save(UC_1, file = "Data/UC_1.RData")
# SACTN_daily_temps <- SACTN_daily_v4.2 %>%
#   left_join(site_list[,c(4,13)], by = "index") %>%
#   filter(index %in% upwelling_sites$index)
# save(SACTN_daily_temps, file = "Data/SACTN_daily_temps.RData")
# #
# SACTN_US1 <- SACTN_daily_temps %>%
#   separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
#   dplyr::rename(insitu_temp = temp)
# save(SACTN_US1, file = "Data/SACTN_US1.RData")

# Upwelling Cell 2:
# site_list_sub_UC <- site_list %>%
#   filter(coast == "wc") %>%
#   filter(length > 3650)
#
# upwelling_sites_UC <- site_list_sub_UC[-1:-8,]
# UC_2 <- upwelling_sites_UC[c(-2,-5),] # New sites falling within the upwelling cell before yzerfontein
# save(UC_2, file = "Data/UC_2.RData")
#
#
# SACTN_daily_temps_UC <- SACTN_daily_v4.2 %>%
#   left_join(site_list[,c(4,13)], by = "index") %>%
#   filter(index %in% upwelling_sites_UC$index)
# save(SACTN_daily_temps, file = "Data/SACTN_daily_temps.RData")
#
# SACTN_US2 <- SACTN_daily_temps_UC %>%
#   separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
#   dplyr::rename(insitu_temp = temp)
#
# save(SACTN_US2, file = "Data/SACTN_US2.RData")

# combined_US <- rbind(UC_1,UC_2)
# save(combined_US, file = "Data/combined_US.RData")
load("~/Documents/Masters_2019/MastersProject/Data/combined_US.RData")
metaData2 <- combined_US
```

## Calculate mean meta data for each coast and add to metaData2
# Dictate coastal groupings

wc <- c("Hout Bay", "Kommetjie", "Port Nolloth", "Sea Point", "Lamberts Bay", "Saldanha Bay", "Yzerfontein", "Dassen Island", "Oudekraal")

```{r}
west <- metaData2[metaData2$site %in% wc,]
west$coast <- "west"
west2 <- data.frame(site = "west coast", coast = "west", lon = round(mean(west$lon), 4),
                    lat = round(mean(west$lat), 4), depth = round(mean(west$depth), 1),
                   type = NA, 'start date' = min(west$`start date`),'end date' = max(west$`end date`), 
                   length = round(mean(west$length), 0), 'temp days' = round(mean(west$'temp days'), 0), 
                   'NA days' = round(mean(west$'NA days'), 0), 'NA%' = round(mean(west$`NA%`),1), 
                   mean = round(mean(west$mean),1), sd = round(mean(west$sd),1), 
                   min = round(mean(west$min),1), max = round(mean(west$max),1))

names(west2)[c(7:8,10:12)] <- c("start date", "end date", "temp days", "NA days", "NA%")
metaData2 <- data.frame(ID = seq(1, length(metaData2$site),1), metaData2)
metaData2$lon <- round(metaData2$lon,2); metaData2$lat <- round(metaData2$lat,2)
metaData2$depth <- NULL; metaData2$NA.days <- NULL

save(metaData2, file = "Data/metaData2.Rdata")
```

# Creating a quick visualisation

```{r}
load("Data/south_africa_coast.RData")
load("Data/sa_provinces_new.RData")

# Setting the plotting parameters
sa_lats <- c(-35.5, -26); sa_lons <- c(14, 34)

sa_site <- ggplot() + 
  geom_polygon(data = south_africa_coast, aes(x = long, y = lat, group = group), 
               show_guide = FALSE, fill = "grey80") +
  geom_path(data = sa_provinces_new, aes(x = long, y = lat, group = group), colour = "grey50") +
  geom_point(data = metaData2, aes(x = lon, y = lat, colour = metaData2$coast, 
                                   shape = metaData2$type), alpha = 0.8, size = 2.2) +
  guides(colour = guide_legend(title = "coast"),
         shape = guide_legend(title = "type"))

ggsave("Figures/map01.pdf", width = 7.5, height = 4.5, pointsize = 10)
```