### 2019-02-26 Chlorophyll_a .Rmd
chloro

---
title: "Chlorophylla"
author: "Amieroh Abrahams"
date: "18 February 2019"
output: html_document
---

The aim of this script is to extract the chlorophyll a netCDF files and convert it to CSV. The chlorophylla data was obtained from MODIS Aqua. The aim of working with this chlorophyll a data is to analyse how it changes/varies with upwelling.

```{r}
library(ncdf4)
library(data.table)
library(tidyverse)
library(reshape2)
library(plyr)
library(lubridate)
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(ggpubr)
library(fasttime)
library(FNN)

MODIS_chlor.dir <- "/home/amieroh/Documents/Data/Datasets/Chlorophyll_a"
MODIS_chlor.csv.dir <- "/home/amieroh/Documents/Data/Datasets"

region <- "BC" # Benguela Current
coords <- bbox[, region]
coords <- c(-35, -20, 10, 20) # this is the BC

ncList <- list.files(path = MODIS_chlor.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
strt.date <- str_sub(basename(ncList[1]), start = 2, end = 8)
end.date <- str_sub(basename(ncList[length(ncList)]), start = 2, end = 8)
nc.init <- nc_open(ncList[1])
LatIdx <- which(nc.init$dim$lat$vals > coords[1] & nc.init$dim$lat$vals < coords[2])
LonIdx <- which(nc.init$dim$lon$vals > coords[3] & nc.init$dim$lon$vals < coords[4])
nc_close(nc.init)

ncFun <- function(nc.file = nc.files, csv.dir = csv.dir) {
nc <- nc_open(nc.file)
  instrument <- ncatt_get(nc, 0, "instrument")$value
  platform <- ncatt_get(nc, 0, "platform")$value
  product_name <- ncatt_get(nc, 0, "product_name")$value
  fNameStem <- substr(product_name, 17, 38)
  timeStamp <- substr(product_name, 2, 8)
  origin <- paste0(substr(timeStamp, 1, 4), "-01-01")
  date <- as.Date(as.numeric(substr(timeStamp, 5, 7)), origin)
  chl <- round(ncvar_get(nc,
                   varid = "chlor_a",
                   start = c(LonIdx[1], LatIdx[1]),
                   count = c(length(LonIdx), length(LatIdx))),
               3)
  dimnames(chl) <- list(lon = nc$dim$lon$vals[LonIdx],
                        lat =  nc$dim$lat$vals[LatIdx])
  nc_close(nc)
  chl <-
    as.data.table(melt(chl, value.name = "chl"), row.names = NULL) %>%
    mutate(t = date) %>%
    na.omit()
  fwrite(chl,
         file = paste(csv.dir, "/", region, "-", instrument, ".",platform, ".",
                      fNameStem, "-", strt.date, "-", end.date, ".csv", sep = ""),
         append = TRUE, col.names = FALSE)
  rm(chl)
}

llply(ncList, ncFun, csv.dir = MODIS_chlor.csv.dir, .parallel = TRUE)
```

## Setting up the chlorophylla dataset

```{r}
MODIS_Chloro <- "~/Documents/Masters_2019/MastersProject/Data"
MODIS_Chloro <- fread(paste0(MODIS_Chloro, "/BC-MODIS.Aqua.L3m_8D_CHL_chlor_a_9km-2002185-2018345.csv"),
            col.names = c("lon", "lat", "chloro", "date"))

# save(MODIS_Chloro, file = "Data/MODIS_Chloro.RData")
load("Data/MODIS_Chloro.RData")
load("Data/combined_US.RData")

daily_chloro_data <- MODIS_Chloro %>%
  dplyr::mutate(date = as.Date(date)) %>% 
  dplyr::group_by(lon, lat, date) %>% 
  mutate(month = month(date),
         day = day(date),
         year = year(date)) %>% 
  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE))

# fastDate <- function(x, tz = NULL) {
#   as.Date(fastPOSIXct(x, tz = tz))
# }
# 
# daily2monthly <- function(dailyData) {
#   monthlyData <- dailyData %>%
#     dplyr::mutate(time = fastDate(time)) %>%
#     dplyr::mutate(month = floor_date(time, unit = "month")) %>%
#     dplyr::group_by(lon, lat, month) %>%
#     dplyr::summarise(chlor_a = mean(chlor_a, na.rm = TRUE)) %>%
#     dplyr::mutate(year = year(month)) %>%
#     dplyr::group_by(lon, lat) %>%
#     dplyr::mutate(num = seq(1:length(chlor_a))) %>%
#     dplyr::ungroup()
#   return(monthlyData)
# }
#  daily2monthly(MODIS_Chloro)

# Pixels and matching
# match_index <- knnx.index(data = as.matrix(daily_chloro[,4:5]),
#                           query = as.matrix(combined_US[,5:6]), k = 1)
# 
# # # Select SST pixels nearest to insitu sites
# pixel_match <- daily_chloro[match_index,] %>%
#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
#   unite(col = day, year, month, sep = "-", remove = F) %>% 
#   mutate(site = combined_US$site)
# 
# daily_chloro_match <- daily_chloro %>%
#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
#   unite(col = day, year, month, sep = "-", remove = F) %>% 
#   filter(combi %in% pixel_match$combi)
# 
# daily_chloro_match_sites <- daily_chloro_match %>%
#   left_join(pixel_match, by = c("combi", "lon", "lat")) %>%
#   group_by(day, site) %>% 
#   dplyr::summarise(chloro = mean(chloro, na.rm = TRUE))
```