### 2019-03-02 wavelets.Rmd

---
title: "Wavelets"
author: "Amieroh Abrahams"
date: "09 February 2019"
output: html_document
---

Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events varies per sites along the west coast of South Africa. It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions. Here i was only required to plot a time series of 2 years.

# Libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
```

### Load site list SACTN data
Now to get to the data. The first step involves the loading of the site list. The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures. These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents. Here we will only focus on the temperature data found along the west coast (wc) (i.e. sites influenced by the Benguela current, EBUS).

```{r}
load("Data/site_list_v4.2.RData")
load("Data/SACTN_daily_v4.2.RData")
```

## Load in the rest of the datasets

* See the Dectect_MCS.Rmd script for more info. The datasets being loaded here are only for temperatures collected between the years 1992-2016. However if a longer time series is required changes may  be made to the Dectect_MCS.Rmd script (line 136)

```{r}
load("Data/overlap_ts.RData")
load("Data/insitu_MUR.RData")
load("Data/insitu_OISST.RData")
load("Data/insitu_G1SST.RData")
load("Data/insitu_K10.RData")
```

# Time series of one year

Here I only look at upwelling duration over a period of 1 year. from 2014 - 2015

```{r}
filtered_years <- function(df){
  upwelling<- df %>% 
  filter(year(date) %in% seq(2014, 2015)) %>% 
  drop_na()
}

SACTN_fyears <- filtered_years(df = overlap_ts) 
MUR_fyears <- filtered_years(df = insitu_MUR) 
OISST_fyears <- filtered_years(df = insitu_OISST) 
G1SST_fyears <- filtered_years(df = insitu_G1SST) 
K10_fyears <- filtered_years(df = insitu_K10)

# save(SACTN_fyears, file = "Data/temp_2014_2015/SACTN_fyears.RData")
# save(MUR_fyears, file = "Data/temp_2014_2015/MUR_fyears.RData")
# save(OISST_fyears, file = "Data/temp_2014_2015/OISST_fyears.RData")
# save(G1SST_fyears, file = "Data/temp_2014_2015/G1SST_fyears.RData")
# save(K10_fyears, file = "Data/temp_2014_2015/K10_fyears.RData")
```

# Wavelet analyses

```{r}

# When doing a wavelet analyses on the SACTN dataset the "insitu_temp" column needed to be renamed to "temp" for the rest of the code to run. The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)
SACTN_fyears <- SACTN_fyears %>%
  dplyr::rename(temp = insitu_temp)

temp.d <- function(df){
temp.d <- df %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, temp, date)
}

temp.d <- temp.d(df = SACTN_fyears) 
# temp.d <- temp.d(df = MUR_fyears) 
# temp.d <- temp.d(df = OISST_fyears)
# temp.d <- temp.d(df = G1SST_fyears) 
# temp.d <- temp.d(df = K10_fyears)

temp.d <- temp.d %>% 
  select(site, no, temp, date)

prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))

ggplot(PN_prewhite, aes(x = no, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()

wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}

PN_wave <- dlply(PN_prewhite, .(site), wl.fun)

for (i in 1:length(PN_wave)) {
  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures/new",plot_name,"_wavelets.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(PN_wave, plot_fun)
```

## Examining seasonal upwelling

Creating a seasonal column. Allowing me to compare the intensity/duration of upwelling on a seasonal basis. This should to be repeated for each of the sites within each of the datasets (SACTN, MUR, OISST, G1SST and K10). The code bellow needs to be edited. Here I only focused on Port Nolloth within the SACTN dataset.

```{r}
load("Data/temp_2014_2015/G1SST_fyears.RData")
load("Data/temp_2014_2015/K10_fyears.RData")
load("Data/temp_2014_2015/MUR_fyears.RData")
load("Data/temp_2014_2015/OISST_fyears.RData")
load("Data/temp_2014_2015/SACTN_fyears.RData")

seasonal_col <- function(df){
  temp_data <- df %>% 
    mutate(month = month(date, abbr = T, label = T),
           year = year(date)) %>% 
    group_by(site, month, year) %>% 
     mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
  return(temp_data)
}

SACTN_seasonal <- seasonal_col(df = SACTN_fyears)
MUR_seasonal <- seasonal_col(df = MUR_fyears)
OISST_seasonal <- seasonal_col(df = OISST_fyears)
G1SST_seasonal <- seasonal_col(df = G1SST_fyears)
K10_seasonal <- seasonal_col(df = K10_fyears)

# save(SACTN_seasonal, file = "Data/temp_2014_2015/SACTN_seasonal.RData")
# save(MUR_seasonal, file = "Data/temp_2014_2015/MUR_seasonal.RData")
# save(OISST_seasonal, file = "Data/temp_2014_2015/OISST_seasonal.RData")
# save(G1SST_seasonal, file = "Data/temp_2014_2015/G1SST_seasonal.RData")
# save(K10_seasonal, file = "Data/temp_2014_2015/K10_seasonal.RData")

load("Data/temp_2014_2015/SACTN_seasonal.RData")
load("Data/temp_2014_2015/MUR_seasonal.RData")
load("Data/temp_2014_2015/OISST_seasonal.RData")
load("Data/temp_2014_2015/G1SST_seasonal.RData")
load("Data/temp_2014_2015/K10_seasonal.RData")
```

A seasonal plotting representing the sea surface temperature for 2014 and 2015

```{r}
PN_14 <- SACTN_seasonal %>% 
  filter(site == "Port Nolloth") %>% 
  filter(year == 2014 )

PN_15 <- SACTN_seasonal %>% 
  filter(year == 2015 )
  # filter(site == "Port Nolloth") %>% 
  # filter(year == 2015 )

combined_seasonal <- function(df){
  plot <- df %>% 
    ggplot(aes(x = date, y = insitu_temp)) +
    geom_line() +
    facet_wrap(~season, ncol = 2, nrow = 2) +
    theme_bw()
}

combined_seasonal_2014 <- combined_seasonal(df = PN_14)
combined_seasonal_2015 <- combined_seasonal(df = PN_15)
combined_PortNolloth <- ggarrange(combined_seasonal_2014, combined_seasonal_2015)

# ggsave(plot = combined_PortNolloth, filename = "Figures/combined_PortNolloth.png")
```

Creating a wavelet analyses for the year 2015. This code creates a plot per season.

```{r}
seasonal_PN_15 <- PN_15 %>% 
  dplyr::rename(temp = insitu_temp)

temp.d <- function(df){
temp.d <- df %>% 
  dplyr::group_by(site, date, season) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(season, site, no, temp)
}

temp.d <- temp.d(df = seasonal_PN_15)

temp.d <- temp.d %>% 
  select(season, site, no, temp)

prewhite_fun <- function(x) {
  df <- x[, 3:4]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site, season), prewhite_fun))

temp_seasonal <- ggplot(temp.d, aes(x = date, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ season, nrow = 4) +
  theme_bw()

wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}

PN_wave <- dlply(temp.d, .(site, season), wl.fun)

for (i in 1:length(PN_wave)) {
  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures/new",plot_name,"_wavelets.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(PN_wave, plot_fun)
```