### 2019-05-09 upwelling_percentile.Rmd

---
title: "upwelling_percentile"
author: "Amieroh Abrahams"
date: "08 May 2019"
output: html_document
---

# Upwelling points

* After analysing the data within the `upwelling_threshold.Rmd` script and plotting the data it is clear that upwelling occurs at a temperature less than or equal (<=) to 12.
* It is also evident that upwelling lasts for a total of 2-3 days

# Aims of this script:
  - Find a 30year time series
  - Then find the count of the number of temperatures found less than or equal to (<=) 12
  - Thereafter, divide that answer by the total number of number temperatures
  - Then we have the percentile

# Questions: How does the intensity and duration of an upwelling event vary at different distances from the coastline??? 
  # Here we can make use of the heatwaveR package to determine the count/durartion and intensity.
  # Got the various pixels : load(Data_P1/site_pixels.RData)
  # Calculate event count, duration and mean intensity once percentile has been determined

# Tips:
  - Look at the percentile level that seawater temperature must decrease below when an upwelling event occurs
  - To what extent the temperature of upwelled water  is colder than the mean SST. 
  - How the minumum temperature of upwelled water becomes to be that temperature
  - Why temperature does not become any colde

 # DETERMINING THE UPWELLING THRESHOLD
  # Plot the data
  # Find a value showing a clear drop in temperature
  # Determine from the dataset the amount of values falling bellow this defined temperature
  # Convert this to percentile
  # Determine how long (in days) upwelling events usually last 
  # Once we have the percentile we can make use of the heatwaveR package to determine the count, intensity, duration, etc.
  # Use one year of data maybe occuring in 1990s and one year of current data to observe how upwelling varied in past and present

# Loading packages

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
library(heatwaveR)
```

# Loading the data

```{r load_files1, include=TRUE}
load("Data_P1/site_list_v4.2.RData")
load("Data_P1/SACTN_daily_v4.2.RData")
```

Finding a time series of greater than 30years

```{r}
site_list_sub <- site_list %>% 
  filter(coast == "wc") %>% 
  filter(length > 10950)

SACTN_temps_US <- SACTN_daily_v4.2 %>% 
  left_join(site_list[,c(4,13)], by = "index") %>% 
  filter(index %in% site_list_sub$index)
# save(SACTN_temps_US, file = "Data/SACTN_temps_US.RData")

SACTN_split <- SACTN_temps_US %>% 
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>% 
  drop_na()

SACTN_temp <- SACTN_split %>% 
  filter(temp <= 13)

percentile <- 30840/60189*100 # Count values
# Percentile value of 50

###################################################Percentile classification#################################################
# percentile = 70
0.7*60189
# 42132
SACTN_TRY1 <- SACTN_split %>%
  arrange(temp) %>% 
  slice(1:42132) # Upwelling temp  = 14 degrees

# Percentile = 60
0.6*60189
# 36113
SACTN_TRY2 <- SACTN_split %>%
  arrange(temp) %>% 
  slice(1:36113) # Upwelling temp = 13.7 degrees

# Percentile = 50
0.5*60189
# 30094
SACTN_TRY3 <- SACTN_split %>%
  arrange(temp) %>% 
  slice(1:30094) # Upwelling temp = 13 degrees

# Percentile = 40
0.4*60189
# 24076
SACTN_TRY4 <- SACTN_split %>%
  arrange(temp) %>% 
  slice(1:24076) # Upwelling temp = 13 degrees

############################################################################################################################

SACTN_PN <- SACTN_split %>%
  filter(site == "Port Nolloth") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_LB <- SACTN_split %>%
  filter(site == "Lamberts Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_SB <- SACTN_split %>%
  filter(site == "Saldanha Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_SP <- SACTN_split %>%
  filter(site == "Sea Point") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_HB <- SACTN_split %>%
  filter(site == "Hout Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()
```

```{r}
ts <- ts2clm(SACTN_PN, climatologyPeriod = c("1974-01-01", "2016-12-31"),
             pctile = 50)
res <- detect_event(ts)
upwelling <- res$climatology # Show a portion of the climatology
upwelling_events <- res$event[1:5,1:10] # Show some of the cold-spells
# MCS <- MCS[14408:15503,] # Get the years from 2013 - 2015

upwelling_SACTN_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2010-01-01", end_date = "2015-12-31")
```

# MUR

```{r}
load("~/Documents/Masters_2019/MastersProject/Exploring/Data/MUR_SST.RData")
MUR_SST_PN <- MUR_SST %>%
  filter(site == "Port Nolloth") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST_PN, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 30)
res <- detect_event(ts)
upwelling <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2011-12-31", end_date = "2014-10-30") +
  ggtitle("MUR_PN") 
```